\documentclass[10pt,lineno]{../_configs}

\articletype{Theoretical} % article type

\bibliography{SFC.bib}

\newcommand{\autor}{\textcite{botta_when_2019} }

\title{
\large{SFC}\vspace{2pt}\\
\Huge{\autor - \citetitle*{botta_when_2019}}
}
\date{April, 21th 2020}

\author[$\ast$]{Gabriel Petrini}

\affil[$\ast$]{PhD Student at Unicamp.}

\keywords{
	AB-SFC model\\
	financial complexity\\
	securitization
}

\runningtitle{Sillabus} % For use in the footer 

%% For the footnote.
\runningauthor{Petrini}

\begin{abstract}
\abssection{Background} Widespread attention to system and financial product complexity after the Great Recession.

\abssection{Supporting ideas} Defines financial complexity as both increase of financial instruments to real-side economics and the intrinsic complexity of financial products created by financial institutions.

\abssection{Aim} Analyze the macroeconomic effects of financial complexity on economic growth, macroeconomic stability and inequality

\abssection{Contribution} 

\abssection{Relevance} Maps some complexity-financial strands. Endogenous boom-and-bust credit cycles.

\abssection{Methods} ABM-SFC approach with heterogeneous household sector. Includes positive feedbacks between households, productive non-financial firms, and financial institutions giving rise to endogenous boom-and-bust credit cycles. Propose an stage-approach in which model complexity increases gradually. \textbf{Commercial banks:} issue credit-money to households and non-financial firms. \textbf{Investment funds:} collects funds from households and allocates between financial assets available

\abssection{Results} Higher financial complexity leads to higher macroeconomic instability, income and wealth inequality and a more crisis-prone economic system

\abssection{Interesting findings} Proposes a SFC model with only one heterogeneous sector (households).

\abssection{Critics} 
 \end{abstract}

\begin{document}

\maketitle
\articletypemark
\marginmark
\thispagestyle{firststyle}
% Please add here a significance statement to explain the relevance of your work

\noindent \textbf{Citation:} 	\fullcite{botta_when_2019}

\begin{infobox}
	\textbf{5SS:} \autor 
\end{infobox}

\begin{redbox}{Further readings}
	\begin{itemize}
		\item 
	\end{itemize}
\end{redbox}

\printbibliography
	
\end{document}
