% Created 2021-01-21 qui 12:28
% Intended LaTeX compiler: pdflatex
\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[top=3cm, bottom=2cm, left=3cm, right=2cm]{geometry}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[theorems, skins]{tcolorbox}
\usepackage[style=abnt,noslsn,extrayear,uniquename=init,giveninits,justify,sccite,
scbib,repeattitles,doi=false,isbn=false,url=false,maxcitenames=2,
natbib=true,backend=biber]{biblatex}
\usepackage{url}
\usepackage[cache=false]{minted}
\usepackage[linktocpage,pdfstartview=FitH,colorlinks,
linkcolor=blue,anchorcolor=blue,
citecolor=blue,filecolor=blue,menucolor=blue,urlcolor=blue]{hyperref}
\usepackage{attachfile}
\usepackage{setspace}
\usepackage{tikz}
\usepackage[document]{ragged2e}
\usepackage[none]{hyphenat} % Avoid hyphenization
\usepackage[right=3cm,left=3cm, footskip=0.25in]{geometry}
\addbibresource{refs.bib}
\usepackage{svg, caption, multirow, booktabs, setspace, indentfirst, authblk}
\usepackage[english]{babel}
\author[1]{Lucas Teixeira}
\affil[1]{Assistent Professor at University of Campinas (Brazil), Email: \url{lucastei@unicamp.br}} % Author affiliation
\author[2]{Gabriel Petrini}
\affil[2]{PhD Student at University of Campinas (Brazil), Email: \url{gpetrinidasilveira@gmail.com}} % Author affiliation
\date{\today}
\title{Long-run effective demand and residential investment: a Sraffian supermultiplier based analysis}
\begin{document}

\maketitle
\begin{abstract}
In this paper, we build a fully specified parsimonious Sraffian supermultiplier stock-flow consistent model (SSM-SFC) with two non-capacity creating autonomous expenditures: residential investment and debt-financed consumption.
Our model represents a closed economy without government with workers and capitalists households and only the latter are not credit constrained.
The introduction of residential investment implies that our SSM-SFC model has two real assets: firms' productive capital and households' real estate.
In our model, residential investment growth rate responds to changes of house price inflation.
The numerical simulation experiments shows that our model keeps the main standard Sraffian supermultiplier growth models results.
As a particular result, an increase of residential investment growth rate implies a decrease of real estate share in total real assets.
Inspired by the recent U.S. bubble episode, we plug house price inflation data (1992-2019) into our model.
Although simple, our numerical simulations are capable to reproduce some stylized facts such as residential investment leading the business cycle and capital accumulation and a clockwise pattern between non-capacity creating autonomous expenditures and capacity utilization rate.

\noindent \textbf{Keywords:} Residential Investment; Sraffian supermultiplier; Asset bubble;  Stock-Flow Consistent approach.
\end{abstract}

\doublespacing
\setlength{\parindent}{4em}
\setlength{\parskip}{1em}
\end{document}
