\documentclass[10pt,lineno]{../_configs}

\articletype{Empirical} % article type

\bibliography{HOUSING.bib}

\newcommand{\autor}{\textcite{sanchez_price_2011} }

\title{
\large{Housing}\vspace{2pt}\\
\Huge{\autor - The Price Responsiveness of Housing Supply in OECD Countries}
}
\date{Date}

\author[$\ast$]{Gabriel Petrini}

\affil[$\ast$]{PhD Student at Unicamp.}

\keywords{
	Keyword1\\ 
	Keyword2\\
	Keyword3
}

\runningtitle{Sillabus} % For use in the footer 

%% For the footnote.
\runningauthor{Petrini}

\begin{abstract}
\abssection{Background} Responsiveness of housing supply to changes in prices bears important implications for the evolution of housing prices and the speed of adjustment of housing markets

\abssection{Supporting ideas} Differences in supply responsiveness to prices implies if housing market responds to demand side shocks with more construction or higher prices. Existing evidence suggests that adjustment occurs in the \textbf{price} of housing rather than in expanding housing supply. An above unite long-run supply elasticity imply a \textbf{housing output increase} relatively more than prices as a result of demand shock.

\abssection{Contribution} Estimates long-run price elasticity of new housing supply.Compares the estimates of the responsiveness of housing supply with those of other studies


\abssection{Methods} Stock-flow model \cite{dipasquale_housing_1994} within an error correction framework for 21 OECD countries. \textbf{Price equation variables:} real income (expected and verified to be positive); real interest rate (user cost variable, \textit{i.e.} opportunity cost of capital, expected and verified  to be negative); stock of dwellings (expected and verified  to be negative); demographic variable (who are more likely to buy housing, expected and verified  to be positive). \textbf{Residential investment equations:}  real residential construction costs (expected and not verified negative for some cases); real house prices (expected and verified  positive); and the same demographic variable (expected and verified positive for some cases). \textbf{Controls}: quarterly dummies for  seasonal effects. ADF tests were employed to check for  the presence of unit roots

\abssection{Results} There is a long-run relationship between real house
prices or investment and the explanatory variables;
New housing supply is relatively more flexible in North
America and some Nordic countries, while it is more rigid in continental European countries and in the
United Kingdom;
The responsiveness of housing supply depends not only on national geographical and
urban characteristics but also on \textbf{policies}, such as land use and planning regulations. 

\abssection{Interesting findings} USA and Nordic countries are more flexible compared to others; Constraints in the supply of housing may alter local employment and wage dynamics across residential areas with repercussions for labor mobility; In flexible-
supply countries, housing investment adjusts more rapidly to large changes in demand and contributes to \textbf{more cyclical swings} in economic growth.

\abssection{Relevance} Housing supply responsiveness to price changes varies substantially across countries. This suggest that countries particularities are relevant. Also shows the relevance of policies, so macroeconomic institution matter. 
Housing supply responsiveness shed light on housing policy reforms aimed at dampening housing price volatility and, in interaction with macro policies, increase macroeconomic resilience to shocks.


\abssection{Critics} Authors uses real interest rate instead mortgage interest rate.
 \end{abstract}

\begin{document}

\maketitle
\articletypemark
\marginmark
\thispagestyle{firststyle}
% Please add here a significance statement to explain the relevance of your work

\noindent \textbf{Citation:} 	\fullcite{sanchez_price_2011}

\begin{infobox}
	\textbf{5SS:} \autor estimates a Stock-Flow model to evaluate house supply responsiveness. Evidence suggest that less responsive house supply leads to an increase in housing prices due to demand shock. However, more responsiveness is also related to a bigger cyclical swings. The authors still argues that a more responsive house supply is more desirable. The estimation also reports a country-specific responsiveness and the relevance of other macro policies.
\end{infobox}

\begin{redbox}{Further readings}
	\begin{itemize}
		\item None
	\end{itemize}
\end{redbox}

\printbibliography
	
\end{document}